### 2022-05-01 outputs/paper/paper.Rmd
Added complete results section with analysis of Granger tests

---
title: "Analysing the effects of COVID-19 on Crime in the United States"
subtitle: "A deep dive into criminal activity during the pandemic"
author: 
  - Nayan Saxena
thanks: "Code and data are available at: https://github.com/the-infiltrator/COVID-19-Crime"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The COVID-19 pandemic caused the various state-level governments in the United States to issue stay-at-home orders in early 2020. These policy measures resulted in sweeping impacts on the everyday life of people with mostly negative implications.  The more negative implications of the pandemic have generally eclipsed the more positive aspects such as the reduction of crime rates by 37% worldwide [@boman2021global], and also the noticeable drop in carbon emissions [@bauwens2020impact] and elevation of air quality around the world—in turn reducing pollution-related respiratory issues  [@dutheil2020covid]. While crime rates globally have gone down, it is also important to take a granular look at the crime data to observe which types of crimes have been most affected,
and further, examine if there is a significant correlation between crime rates and COVID-19 cases — which is precisely the goal of this study. We achieve our goal of analyzing the pandemic data in conjunction with crime data obtained from the official city websites through time series modeling and also exploratory data analysis focusing on the criminal activity in key US cities such as Chicago,  Los Angeles, Philadelphia, and Seattle.  Each city has been chosen specifically to reflect different American aspects, geographical as well as socio-economic, to enable us to form more generalizable as well as granular conclusions. \\par
 \\textbf{Keywords:} COVID-19 Pandemic, Criminal Behaviour, Time Series Modelling, Correlation, Dataset Creation, Criminal Psychology"

output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
urlcolor: blue

pdf_document:
  extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(plyr)
library(flexdashboard)
library(RSocrata)
library(tidyverse)
library(tsbox) # transform data into time series
library(xts)
library(COVID19) # to get data about covid 19
library(forecast) #ariRI model
library(vars) #VAR and Causality
library(dygraphs)
library(leaflet)
library(htmlwidgets)
#[INCOMPLETE] Graphs in Chicago have been converted here to Plotly HTML Widgets 
# Make some noisily increasing data [Testing Purposes]
set.seed(955)
dat <- data.frame(cond = rep(c("A", "B"), each=10),
                  xvar = 1:20 + rnorm(20,sd=3),
                  yvar = 1:20 + rnorm(20,sd=3))
#### COVID19 DATA ####
#Load Chicago Data
covid19_CH <- covid19("USA", level = 3) |>
  # this cook county contains chicago
  filter(administrative_area_level_3 == "Cook",
         administrative_area_level_2 == "Illinois" ) |>
  # filter out days when confirmed is zero or one
  # becasue when it was 2 for a very long time
  filter(confirmed > 2)

# Load Providence data
covid19_RI <- covid19("USA", level = 2) |>
  filter(administrative_area_level_2 == "Rhode Island") |>
  # filter out days when confirmed is zero or one
  # becasue when it was 1 for a very long time
  filter(confirmed > 1)
## March 07 has 140 confirmed case which is impossible.
## Google shows that date had still 3 cumulative case
## Manual adjustment on row 5
covid19_RI$confirmed[5] = covid19_RI$confirmed[4]

# Load Boston Data
covid19_MA <- covid19("USA", level = 2) |>
  filter(administrative_area_level_2 == "Massachusetts") |>
  # filter out days when confirmed is zero or one
  # becasue when it was 1 for a very long time
  filter(confirmed > 1)

# Load LA data
# extract LA data from US data. 
covid19_LA <- covid19("USA", level = 3) |>
  filter(administrative_area_level_3 == "Los Angeles",
         administrative_area_level_2 == "California") |>
  # stayed at 1 for long time
  filter(confirmed > 1)

# Load Atlanta Data
covid19_AT <- covid19("USA", level = 3) |>
  filter(administrative_area_level_2 == "Georgia",
         administrative_area_level_3 == 'Fulton') |>
  # filter out days when confirmed is zero or one
  # becasue when it was 2 for a very long time
  filter(confirmed > 2)

# Load Seattle Data
covid19_SEA <- covid19("USA", level = 3) |>
  # this cook county contains chicago
  filter(administrative_area_level_3 == "King",
         administrative_area_level_2 == "Washington" ) |>
  # filter out days when confirmed is zero or one
  # becasue when it was 1 for a very long time
  filter(confirmed > 1)

# extract Pennsylvania data from US data. 
covid19_PA <- covid19("USA", level = 3) |>
  filter(administrative_area_level_2 == "Pennsylvania",
         administrative_area_level_3 == "Philadelphia") |>
  # filter out days when confirmed is zero
  filter(confirmed > 0)

chicago<- read.csv2('data/chicago.csv')
la<- read.csv2('data/la.csv')
seattle<- read.csv2('data/seattle.csv')
phil<- read.csv2('data/phil.csv')

# summary of all crimes
chicago_summary <- chicago |>
  group_by(primary_type) |>
  dplyr::summarise(number_of_crime = dplyr::n()) |>
  arrange(desc(number_of_crime))

phil_summary <- phil |>
  group_by(text_general_code) |>
  summarise(number_of_crime = n()) |>
  arrange(desc(number_of_crime))

LA_summary <- la |>
  group_by(crm_cd_desc) |>
  summarise(number_of_crime = n()) |>
  arrange(desc(number_of_crime))

seattle_summary <- seattle |>
  group_by(offense) |>
  dplyr::summarise(number_of_crime = dplyr::n()) |>
  arrange(desc(number_of_crime))
```

# Introduction
The COVID-19 pandemic caused the various state-level governments in the United States to issue stay-at-home orders in early 2020. These policy measures resulted in sweeping impacts on the everyday life of people with mostly negative implications.  The onset of the pandemic brought with it a host of other issues like job losses, unemployment, financial crisis, and mental health issues. The pandemic also forced people to stay at home for prolonged periods, sometimes over a month, which seemed to have exasperated opioid abuse and also relapse of a host of other addictions. Furthermore, it has also been shown that the pandemic has led to increased stress and anxiety levels among individuals [@boman2020has].  These negative implications of the pandemic have generally eclipsed the more positive aspects such as the reduction of crime rates by 37% worldwide [@boman2021global], and also the noticeable drop in carbon emissions [@bauwens2020impact] and elevation of air quality around the world—in turn reducing pollution-related respiratory issues  [@dutheil2020covid]. While crime rates globally have gone down, it is also important to take a granular look at the crime data to observe which types of crimes have been most affected,
and further, examine if there is a significant correlation between crime rates and COVID-19 cases — which is precisely the goal of this study.

The main goal of our paper is to examine the effect that the pandemic has brought in terms of specific crimes such as assault, battery, criminal theft, and much more based on state-level crime data within the United States. The main motivation behind this research is the scarcity of work examining the effect of the pandemic on crime, even though there has been a lot of work done examining other aspects concerning the pandemic. News stations have consistently reported lower crime rates but have based their analyses on simply the rate of 911 calls and not actual police data. We aim to bridge this gap by focusing on the criminal activity in key US cities such as Chicago,  Los Angeles, Philadelphia, and Seattle.  Each city has been chosen specifically to reflect different American aspects, geographical as well as socio-economic, to enable us to form more generalizable as well as granular conclusions.

The key motivation for our work is previous work showing that certain crimes like domestic violence saw a significantly smaller drop as compared to other crimes after the onset of the pandemic [@bullinger2020covid], indicating the importance of the nature of the actual crime committed. Our focus on individual city data is due to known data collection issues for such research [@boman2020has] and is motivated by preliminary studies showing that the crime rates for specific crimes are largely dependent on the city [@ashby2020initial]. There are also other factors such as the George Floyd protests in June 2020 that we hypothesize may have resulted in a noticeable increase in the vandalism and theft crimes in certain cities.

Overall, our paper is a small contribution building upon and motivated by several recent papers that have covered the impact of the pandemic on the human psyche and crime rates across different states and countries [@boman2021global, @campedelli2020disentangling, @nivette2021global] alongside work prompting further research in the area [@eisner2020violence]. We achieve our goal of analyzing the pandemic data in conjunction with crime data obtained from the official city websites, using R statistical programming [@citeR], through time series modeling and also exploratory data analysis.

In the following section, we outline how data for this study was collected alongside an overview of the data. In Section 3 we outline, in detail our modeling approaches and in Section 4 the results are provided for each city. We conclude our study with a detailed discussion in Section 5, explaining some of our key findings and conclusions.

```{r ,fig.cap = "Geographical overview of the cities being examined under this study.",fig.height = 5, fig.width = 5, out.width = "300px", message = FALSE, echo = FALSE, warning=FALSE, fig.align="center"}
library(mapview)
Location <- c("Los Angeles","Chicago","Seattle","Philadelphia" )
lat <- c(33.82377,41.78798,47.714965, 39.952583)
lng <- c( -118.2668,-87.7738,-122.127166 ,-75.165222)
df <- data.frame(Location, lat,lng)

m <- leaflet() |>
  addTiles() |>  # Add default OpenStreetMap map tiles
  addMarkers(data=df,)
mapshot(m, file = "~/Rplot.png")
knitr::include_graphics("~/Rplot.png")
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
```

# Data
The primary data set used is publicly available from the the R, COVID-19 Data Hub [@cdhub] containing a daily summary of COVID-19 cases, deaths, recovered, tests, vaccinations, and hospitalizations for 230+ countries, 760+ regions, and 12000+ administrative divisions of lower level including policy measures, mobility, and geospatial data. Furthermore, we use city data for each city involved in our study to examine the crime rates across the pandemic and the city-wise data sources are outlined below alongside exploratory analyses of both COVID-19 and crime data.

## Chicago
The city of Chicago was chosen for this analysis because it has the highest crime rates in America. The gang culture and gun laws have made Chicago the hub of specifically violent crimes, with higher homicide rates on average as compared to other cities.  Furthermore, during the early stages of the COVID-19 pandemic, it was found that Chicago was one of the first few cities to harbor several virus strains [@anderson], leading to a disaster proclamation by then Governor Pritzker.  Chicago has since then seen roughly three waves of the virus as shown in Figure \@ref(fig:graph1). The data for the city of Chicago was sourced from the [Chicago police](https://home.chicagopolice.org/statistics-data/data-requests/), spanning six main categories of criminal activity which are assault, battery, criminal damage, deceptive practices, vehicle theft, and general theft.

```{r graph1, fig.cap = "Daily Confirmed Cases of COVID-19 in Chicago",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = chicago|>mutate(Date=as.Date(Date))

ts_CH <- covid19_CH %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# try first log difference
ts_diff_CH <- na.omit(diff(ts_CH))
plot.xts(ts_diff_CH,
         main = "Daily confirmed cases of COVID-19 in Chicago")

# extract top 5 crime
top5crime <- chicago %>%
  filter(primary_type %in% head(chicago_summary$primary_type, 5)) %>%
  group_by(Date, primary_type) %>%
  tally() %>%
  spread(primary_type, n)

# rename columns
colnames(top5crime) <- c('time',
                         "assault",
                         "battery",
                         "criminal",
                         'deceptive',
                         "theft")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))

```
When closely examining yearly trends for the different crimes, Figure \@ref(fig:graph2), it can be observed that theft rates and battery noticeably went down in March of 2020 while crimes like assault and Vehicle Theft remained steady. Furthermore, it can be observed throughout the pandemic, these levels remained relatively consistent throughout 2021 and 2022.

```{r graph2, fig.cap = "Yearly Trends for the top five crimes in Chicago",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}
# 2020 only
plt <- copy %>%
  dplyr::select(y_month, month, primary_type, year) %>%
  filter(primary_type %in% chicago_summary$primary_type[1:5] | primary_type == "MOTOR VEHICLE THEFT", y_month != "2020-06") %>%
  dplyr::count(year, month, primary_type) %>%
  na.omit() %>% 
  mutate(Date = as.Date(paste("2000", month, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = year, color = as.character(year))) + 
  geom_line() + 
  facet_wrap(~primary_type, nrow = 1) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
plt
```
In general, it is also noted that most crimes, with the exception of vehicle theft and deception, went up generally in summer and then reduced when the temperatures fell for winter, indicating a seasonal trend. Finally, a more granular look at the frequency of each crime only during the pandemic, as shown in Figure \@ref(fig:graph3), shows that most crimes during the pandemic remained relatively consistent with the exception of criminal damage and theft. A sudden uptick in criminal damage and theft is noted in June 2020, which can be attributed to the violence caused by certain bad actors during the George Floyd Protests.

```{r graph3, fig.cap = "Frequency of each crime in Chicago during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 Onwards only

chicago_daily <- copy |>
  dplyr::select(Date, primary_type, year) |>
  filter(primary_type %in% chicago_summary$primary_type[1:5] | primary_type == "MOTOR VEHICLE THEFT", 
         year >= 2020) |>
 dplyr::count(Date, primary_type) |>
  na.omit() |>
  ggplot(aes(Date, n, group = primary_type, color = primary_type)) +
  geom_line() +
  facet_wrap(~primary_type, nrow = 2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(chicago_summary$primary_type[1:5]))) +
  ylab('Number of Reported Crime') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
chicago_daily
# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))

```

## Los Angeles

```{r graph5, fig.cap = "Daily Confirmed Cases of COVID-19 in L.A",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 3, fig.width = 8,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = la|>mutate(date_occ=as.Date(date_occ))

ts_LA <- covid19_LA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# try first log difference
ts_diff_LA <- na.omit(diff(ts_LA))
plot.xts(ts_diff_LA,
         main = "Daily confirmed cases of COVID-19 in LA")

```
\newpage

Los Angeles, located in California, has historically been a hub of criminal activities mainly centered around narcotics due to its proximity to Tijuana and in general the US-Mexico border. Crime in Los Angeles is also prevalent due to the presence of one of the largest populations of homeless people in the United States mostly within the Skid-Row locality. Los Angeles was at the peak of criminal activity between 1970 and 1990, and more recently in 2015, it was revealed that the Los Angeles Police were under-reporting criminal activity for roughly 8 years, artificially lowering the crime rates and making it an intriguing city to examine from a data perspective [@bird2018impact].

To examine the crime rates within the city we source data directly from the  [Los Angeles Police Department Socrata API](https://dev.socrata.com/foundry/data.lacity.org/2nrs-mtv8) and look at the top five crimes which are mainly variants of burglary and theft crimes, alongside vehicle theft and assault. Finally, as shown in Figure \@ref(fig:graph5), we note that the pandemic had a significant impact in Los Angeles with some of the first patients of COVID-19, including "patient zero", being from Los Angeles. California overall also faced a severe health crisis with cases up to ten times more than were anticipated.

Further examining yearly trends for several crimes, based on Figures \@ref(fig:graph6) and \@ref(fig:graph7), it was found that several crimes like battery/assault, vehicle theft, and burglary increased. The noticeable increase in these crimes remained consistent for years spanning the pandemic, although a noticeable drop in battery, burglary, and vehicle theft was observed in January 2022.

```{r graph6, fig.cap = "Yearly Trends for the top crimes in Los Angeles",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
ty <- copy %>%
  dplyr::select(y_month, month, crm_cd_desc, year) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5],
         y_month != "2020-06") %>%
  count(year, month, crm_cd_desc) %>%
  na.omit() %>%
  mutate(Date = as.Date(paste("2000", month, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n,
             group = year,
             color = as.character(year))) +
  geom_line() +
  facet_free(~crm_cd_desc) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))

ty
```

```{r graph7, fig.cap = "Frequency of each crime in LA during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
daily <- copy %>%
  dplyr::select(date_occ, crm_cd_desc, year) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5],
         year >= 2020) %>%
  count(date_occ, crm_cd_desc) %>%
  ggplot(aes(date_occ, n, group = crm_cd_desc, color = crm_cd_desc)) +
  geom_line() +
  facet_wrap(~crm_cd_desc, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(LA_summary$crm_cd_desc[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))

daily
```
\newpage

## Seattle

```{r graph8, fig.cap = "Daily Confirmed Cases of COVID-19 in Seattle",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = seattle|>mutate(Date=as.Date(Date))

ts_ST <- covid19_SEA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# try first log difference
ts_diff_ST <- na.omit(diff(ts_ST))
plot.xts(ts_diff_ST,
         main = "Daily confirmed cases of COVID-19 in Seattle")
```

```{r graph9, fig.cap = "Yearly Trends for the top crimes in Seattle",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
yty <- copy %>%
  dplyr::select(y_month, MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5],
         y_month != "2020-06") %>%
  dplyr::count(YEAR, MONTH, offense) %>%
  na.omit() %>%
  mutate(Date = as.Date(paste("2000", MONTH, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = YEAR, color = as.character(YEAR))) +
  geom_line() +
  facet_free(~offense, scales = 'free') +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b") +
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

yty
```

One of the fastest-growing cities in the United States, Seattle, was specifically chosen for this study due to its high rates of property crime while having relatively constant overall crime rates across the decades.  As shown in Figure \@ref(fig:graph8), the city was relatively less impacted by COVID-19, making it an interesting candidate in contrast to other cities when it comes to observing criminal behavior. Seattle is also the location of the earliest clinical trial being conducted for the Moderna COVID-19 vaccine which occurred around March 20, 2020 [@tuesday_2020], and more recently has seen a sudden increase in criminal activity alongside increased homelessness [@westneat_2022].

 The crime data for Seattle was sourced from the public [Seattle Police Department Records](https://data.seattle.gov/Public-Safety/SPD-Crime-Data-2008-Present/tazs-3rd5), with our focus mainly being on the top crimes comprising assault, burglary, vandalism, larceny and theft from a vehicle. Given the relatively low rate of COVID-19 in Seattle, trends across most crimes seemed to remain constant before and after the onset of the pandemic as seen in Figure \@ref(fig:graph9). It was also observed that crimes such as burglary increased while theft from motor vehicles showed a decline as compared to previous trends.  These relatively consistent trends during the pandemic are also visible in Figure \@ref(fig:graph10), with an increase in larceny, burglary, and vandalism crimes noted in June-July 2020 due to the Goerge Floyd protests.

```{r graph10, fig.cap = "Frequency of each crime in Seattle during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")

seattle_daily <- copy %>%
  dplyr::select(Date, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR>='2020') %>%
  dplyr::count(Date, offense) %>%
  ggplot(aes(Date, n, group = offense, color = offense)) +
  geom_line() +
  facet_wrap(~offense, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(seattle_summary$offense[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))

seattle_daily
```

## Philadelphia

```{r graph11, fig.cap = "Daily Confirmed Cases of COVID-19 in Philadelphia",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = phil|>mutate(date=as.Date(date))

ts_PH <- covid19_PA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# try first log difference
ts_diff_PH <- na.omit(diff(ts_PH))
plot.xts(ts_diff_PH,
         main = "Daily confirmed cases of COVID-19 in Philadelphia")
```
The last city in our study is Philadelphia, which is one of the most violent cities in America along with the highest poverty rates. The city is the sixth most populous city in the United States with a rich cultural diversity making it ideal for our study. The city was significantly affected by the pandemic as noted by its poor economic performance and decline in the supply of workers [@haider_eichel].  It should be noted though that when compared to the other cities in our study, Philadelphia has the lowest number of daily cases on average as shown in Figure Figure \@ref(fig:graph11).

To examine the impact on crime due to the pandemic we sourced data from the [Philadelphia Police Department](https://www.phillypolice.com/crime-maps-stats/), with an emphasis on the top crimes which were basically assaults, other offenses, thefts from vehicles, general thefts and vandalism. Overall it was found, as demonstrated in Figure \@ref(fig:graph12) that petty offenses increased during the pandemic while most other crimes remained relatively consistent with patterns emerging in prior years.  An immediate downtrend was also observed in crimes such as assaults and thefts right after the onset of the pandemic.

```{r graph12, fig.cap = "Yearly Trends for the top crimes in Philadelphia",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
plt <- phil %>%
  dplyr::select(MONTH, dispatch_date, text_general_code, YEAR, y_month) %>%
  filter(text_general_code %in% phil_summary$text_general_code[1:5]) %>%
  count(YEAR, MONTH, text_general_code) %>%
  na.omit() %>% 
  mutate(Date = as.Date(paste("2000", MONTH, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = YEAR, color = as.character(YEAR))) + 
  geom_line() + 
  facet_wrap(~text_general_code, nrow = 1) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))
plt
```

Further examining the crime trends during years spanning the pandemic, as shown in Figure \@ref(fig:graph13), we observe that there is a general downtrend in petty offenses while most other crimes remain consistent with some seasonal trends in assaults and thefts.  Finally, as noted previously for other cities, an uptick in criminal activity is observed during the black lives matter protests of June 2020.

```{r graph13, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
daily <- copy %>%
  dplyr::select(date, text_general_code, YEAR, y_month) %>%
  filter(text_general_code %in% phil_summary$text_general_code[1:5], YEAR >= 2020) %>%
  count(date, text_general_code) %>%
  na.omit() %>%
  ggplot(aes(date, n, group = text_general_code, color = text_general_code))+
  geom_line() + 
  facet_wrap(~text_general_code, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(phil_summary$text_general_code[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
daily

```

Next, we take a look at the modelling approaches employed to further verify correlations between COVID-19 and crime data.

\newpage

# Methods

## Vector autoregression Model (VAR)

Vector autoregression (VAR) is a statistical model used to capture the relationship between multiple quantities as they change over time. VAR is a type of stochastic process model. VAR models generalize the single-variable (univariate) autoregressive model by allowing for multivariate time series. VAR models are often used in economics and the natural sciences.
A pth-order VAR is denoted "VAR(p)" and sometimes called "a VAR with p lags". A pth-order VAR model is written as

$${\displaystyle y_{t}=c+A_{1}y_{t-1}+A_{2}y_{t-2}+\cdots +A_{p}y_{t-p}+e_{t},\,}y_{t}=c+A_{1}y_{{t-1}}+A_{2}y_{{t-2}}+\cdots +A_{p}y_{{t-p}}+e_{t},\,$$

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc.

## Granger Causality Test

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui. Maecenas velit massa, fringilla vel tristique vitae, condimentum ut purus. Proin commodo eros iaculis, faucibus dui et, vehicula felis. Integer ligula urna, feugiat quis tristique vel, cursus nec quam. Mauris auctor sodales mauris, sed tristique quam sagittis in. Mauris ut bibendum enim. Maecenas consequat elit ut posuere aliquet. In hac habitasse platea dictumst. Integer est arcu, semper ac dolor at, pulvinar vehicula mi.

<!-- \newpage -->

## Forecasting and Impulse Response

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui. Maecenas velit massa, fringilla vel tristique vitae, condimentum ut purus. Proin commodo eros iaculis, faucibus dui et, vehicula felis. Integer ligula urna, feugiat quis tristique vel, cursus nec quam. Mauris auctor sodales mauris, sed tristique quam sagittis in. Mauris ut bibendum enim. Maecenas consequat elit ut posuere aliquet. In hac habitasse platea dictumst. Integer est arcu, semper ac dolor at, pulvinar vehicula mi.

\newpage

# Results 
We now employ the methods outlined in the previous section to narrow down on the crimes that are the most significantly correlated with COVID-19 activity in the respective cities. For each city, the impulse response alongside the model predictions are visualized for the crimes that are most significantly associated with COVID-19.

## Chicago

Based on the fitted VAR models for top crimes in Chicago, we find that COVID-19 cases do not have a significant Granger instantaneous causality, at a 95% confidence level, with crimes such as criminal damage ($p=0.8271$), and battery ($p=0.63$). Whereas, a statistically significant instantaneous correlation was found between COVID-19 cases impacting assault ($p=0.036$) and a significant Granger-causality with the cases impacting vehicle theft ($p = 0.00022$). The impulse response of a new case to these crimes and the model predictions are outlined in Figures \@ref(fig:r11) & \@ref(fig:r12).

```{r top 5 crime VAR, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}

covid19_CH_diff <- data.frame(diff(covid19_CH$confirmed))
colnames(covid19_CH_diff)[1] = "confirmed"
covid19_CH_diff$date = covid19_CH$date[2:length(covid19_CH$date)]
# time as integer
covid19_CH_diff$timeInt = as.numeric(covid19_CH_diff$date)
# make a copy to avoid perfect collinearity
covid19_CH_diff$timeIid = covid19_CH_diff$timeInt

# make a copy to avoid perfect collinearity
covid19_CH_diff$timeIid = covid19_CH_diff$timeInt

gamCH <- gamm4::gamm4(confirmed ~  s(timeInt, k=50), random = ~(1|timeIid), 
	data=covid19_CH_diff, family=poisson(link='log'))

toPredict = data.frame(time = seq(covid19_CH_diff$date[1], 
                                          covid19_CH_diff$date[length(covid19_CH_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# obtain forecast
forecast <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamCH$gam, toPredict, se.fit=TRUE))))

# access residuals
CH_res <- data.frame(covid19_CH_diff$confirmed - forecast$fit)

# transform into time series
CH_res$time = covid19_CH_diff$date
colnames(CH_res)[1] = "residuals"

col_order <- c("time", "residuals")
CH_res <- CH_res[, col_order]

CH_res_ts <- ts_xts(CH_res)

common_time <- seq.Date(start(CH_res_ts), as.Date("2021-12-31"), by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       CH_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])
# plot.xts(CH_res_ts)
# adf test
# adf.test(as.ts(CH_res_ts))
# Stationary process
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
optimal_assault <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_battery <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_criminal <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_deceptive <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_theft <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

VAR_assault <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]), p=optimal_assault$selection[1])
VAR_battery <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]), p=optimal_battery$selection[1])
VAR_criminal <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                    p=optimal_criminal$selection[1])
VAR_deceptive <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                     p=optimal_deceptive$selection[1])
VAR_theft <- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]), p=optimal_theft$selection[1])
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
causality(VAR_assault, cause = colnames(combined_diff)[1])
causality(VAR_assault, cause = "residuals")

causality(VAR_battery, cause = colnames(combined_diff)[2])
causality(VAR_battery, cause = "residuals")

causality(VAR_criminal, cause = colnames(combined_diff)[3])
causality(VAR_criminal, cause = "residuals")

causality(VAR_deceptive, cause = colnames(combined_diff)[4])
causality(VAR_deceptive, cause = "residuals")

causality(VAR_theft, cause = colnames(combined_diff)[5])
causality(VAR_theft, cause = "residuals")

vehicle <- chicago %>%
  filter(primary_type == 'MOTOR VEHICLE THEFT')%>%
  group_by(date, primary_type) %>%
  tally() %>%
  spread(primary_type, n)

colnames(vehicle) <- c('time', 'vehicle')
vehicle_xts <- ts_xts(na.omit(vehicle)[,1:2])
vehicle_diff <- na.omit(diff(vehicle_xts))

combined_diff2 <- merge(vehicle_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       CH_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])
optimal_vehicle <- VARselect(na.omit(combined_diff2)[,c(1,2)], type = 'none', lag.max = 10)
VAR_vehicle <- VAR(y=as.ts(na.omit(combined_diff2)[,c(1,2)]), p=optimal_vehicle$selection[1])
# 
causality(VAR_vehicle, cause = colnames(combined_diff2)[1])
causality(VAR_vehicle, cause = "residuals")
```

```{r r11, fig.show="hold",  fig.cap = "Impact of a new COVID-19 case on assault (left), and VAR model prediction for assault (right).", out.width="50%", message = FALSE, echo = FALSE, warning=FALSE}
# use car theft and criminal damage
par(mfrow = c(1,2))
lags = c(1:16)

# criminal damange
# only significant one is from covid to crime
irf_assault_2 <- irf(VAR_assault, 
                         impulse = "residuals", 
                         response = "assault", 
                         n.ahead = 15,
                         ortho = F)

# ggplot version of it.
irf_assault_2_gg <- data.frame(
  irf_assault_2$irf$residuals[,1],
  irf_assault_2$Lower$residuals[,1],
  irf_assault_2$Upper$residuals[,1]
)

colnames(irf_assault_2_gg) <- c("mean", "lower", "upper")

i1 <- ggplot(irf_assault_2_gg, aes(x = lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  theme_classic() +
  ggtitle("") +
  xlab("Days after a new COVID-19 case")+
  ylab("Assualt")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
i1

f_assault <- forecast::forecast(VAR_assault)

autoplot(f_assault$forecast$assault) +
  ggtitle("Prediction on how many more assault cases compared to yesterday") +
  theme_minimal() +
  ylab("Day-to-day change") +
  xlab(paste("Number of days since", common_time[1]))

```

```{r r12, fig.show="hold",  fig.cap = "Impact of a new COVID-19 case on vehicle thefts (left), and VAR model prediction for vehicle thefts (right).", out.width="50%", message = FALSE, echo = FALSE, warning=FALSE}
# use car theft and criminal damage
par(mfrow = c(1,2))
lags = c(1:16)

# criminal damange
# only significant one is from covid to crime
irf_criminal_2 <- irf(VAR_vehicle, 
                         impulse = "residuals", 
                         response = "vehicle", 
                         n.ahead = 15,
                         ortho = F)

# ggplot version of it.
irf_criminal_2_gg <- data.frame(
  irf_criminal_2$irf$residuals[,1],
  irf_criminal_2$Lower$residuals[,1],
  irf_criminal_2$Upper$residuals[,1]
)

colnames(irf_criminal_2_gg) <- c("mean", "lower", "upper")

i1 <- ggplot(irf_criminal_2_gg, aes(x = lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  theme_classic() +
  ggtitle("") +
  xlab("Days after a new COVID-19 case")+
  ylab("Vehicle Theft")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
i1

interval_value_formatter <- "function(num, opts, seriesName, g, row, col) {
  value = g.getValue(row, col);
  if(value[0] != value[2]) {
    lower = Dygraph.numberValueFormatter(value[0], opts);
    upper = Dygraph.numberValueFormatter(value[2], opts);
    return '[' + lower + ', ' + upper + ']';
  } else {
    return Dygraph.numberValueFormatter(num, opts);
  }
}"

f_criminal <- forecast::forecast(VAR_criminal)

autoplot(f_criminal$forecast$criminal) +
  ggtitle("Prediction on how many more Vehicle Theft cases compared to yesterday") +
  theme_minimal() +
  ylab("Day-to-day change") +
  xlab(paste("Number of days since", common_time[1]))

```

## Los Angeles

When it comes to Los Angeles, only vandalism stood out based on the fitted VAR models. We find that COVID-19 cases do not have a significant Granger instantaneous causality, at a 95% confidence level, with crimes such as battery ($p=0.67$), theft from vehicle ($p=0.9$), vehicle theft ($p=0.81$),  and burglary ($p=0.26$). Whereas, a statistically significant Granger causality was found between COVID-19 cases impacting vandalism ($p=0.003$). The impulse response of a new case to vandalism and the model predictions are outlined in Figure \@ref(fig:r21).

```{r top 5 crime VAR2, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}

LA<-la
top5crime <- LA %>%
  filter(crm_cd_desc %in% head(LA_summary$crm_cd_desc, 5)) %>%
  group_by(date_occ, crm_cd_desc) %>%
  tally() %>%
  spread(crm_cd_desc, n)

# rename columns
colnames(top5crime) <- c('time',
                         "battery",
                         "burglary",
                         'burglary from vehicle',
                         "vandalism",
                         'vehicle')
# create date
top5crime$time <- as.Date(top5crime$time,
                          format = "%Y-%m-%d")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))

ts_LA <- covid19_LA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# try first log difference
ts_diff_LA <- diff(ts_LA)
# still clearly not stationary
# need de-trend

# de-trend 
# GAMM model from STA303 A3

# calculate the difference per day
covid19_LA_diff <- data.frame(diff(covid19_LA$confirmed))
colnames(covid19_LA_diff)[1] = "confirmed"
covid19_LA_diff$date = covid19_LA$date[2:length(covid19_LA$date)]

# time as integer
covid19_LA_diff$timeInt = as.numeric(covid19_LA_diff$date)
# make a copy to avoid perfect collinearity
covid19_LA_diff$timeIid = covid19_LA_diff$timeInt

# make a copy to avoid perfect collinearity
covid19_LA_diff$timeIid = covid19_LA_diff$timeInt

# GAMM model
# 50 too overfit. 15 looks decent
gamLA <- gamm4::gamm4(confirmed ~  s(timeInt, k=90), random = ~(1|timeIid), 
	data=covid19_LA_diff, family=poisson(link='log'))

# obtain fitted value
toPredict = data.frame(time = seq(covid19_LA_diff$date[1], 
                                          covid19_LA_diff$date[length(covid19_LA_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# obtain forecast
forecast <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamLA$gam, toPredict, se.fit=TRUE))))

# access residuals
LA_res <- data.frame(covid19_LA_diff$confirmed - forecast$fit)

# transform into time series
LA_res$time = covid19_LA_diff$date
colnames(LA_res)[1] = "residuals"

col_order <- c("time", "residuals")
LA_res <- LA_res[, col_order]

LA_res_ts <- ts_xts(LA_res)
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
common_time <- seq.Date(start(LA_res_ts), as.Date("2021-12-31"), by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       LA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

# construct VAR
optimal_battery <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_burglary <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_burglary_vehicle <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_vehicle <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

VAR_battery <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]),
                   p=optimal_battery$selection[1])
VAR_burglary <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]),
                    p=optimal_burglary$selection[1])
VAR_burglary_vehicle <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                    p=optimal_burglary_vehicle$selection[1])
VAR_vandalism <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                    p=optimal_vandalism$selection[1])
VAR_vehicle <- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]),
                   p=optimal_vehicle$selection[1])
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
causality(VAR_battery, cause = colnames(combined_diff)[1])
causality(VAR_battery, cause = "residuals")

causality(VAR_burglary, cause = colnames(combined_diff)[2])
causality(VAR_burglary, cause = "residuals")

causality(VAR_burglary_vehicle, cause = colnames(combined_diff)[3])
causality(VAR_burglary_vehicle, cause = "residuals")

causality(VAR_vandalism, cause = colnames(combined_diff)[4])
causality(VAR_vandalism, cause = "residuals")

causality(VAR_vehicle, cause = colnames(combined_diff)[5])
causality(VAR_vehicle, cause = "residuals")

# causality(VAR_vehicle, cause = colnames(combined_diff2)[1])
# causality(VAR_vehicle, cause = "residuals")
```

```{r r21, fig.show="hold",  fig.cap = "Impact of a new COVID-19 case on vandalism (left), and VAR model prediction for vandalism (right).", out.width="50%", message = FALSE, echo = FALSE, warning=FALSE}
# use car theft and criminal damage
lags = c(1:16)

par(mfrow = c(1,2))

# battery
# criem significant to covid
irf_vandalism_1 <- irf(VAR_vandalism,
                     impulse = "residuals",
                     response = "vandalism",
                     n.ahead = 15)

# html plot for vandalism
irf_vandalism_1_gg <- data.frame(
  irf_vandalism_1$irf$residuals[,1],
  irf_vandalism_1$Lower$residuals[,1],
  irf_vandalism_1$Upper$residuals[,1]
)

colnames(irf_vandalism_1_gg) <- c("mean", "lower", "upper")

irf_vandalism_1_plot <- ggplot(irf_vandalism_1_gg, aes(x=lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  ggtitle("") +
  xlab("Days after a new COVID-19 case")+
  ylab("Vandalism")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

irf_vandalism_1_plot

f_vandalism <- forecast::forecast(VAR_vandalism)

autoplot(f_vandalism$forecast$vandalism) +
  ggtitle("Prediction on how many more vandalism cases compared to yesterday") +
  theme_minimal() +
  ylab("Day-to-day change") +
  xlab(paste("Number of days since", common_time[1]))

```

## Philadelphia

Coming to Philadelphia, only vehicle and general thefts stood out based on the fitted VAR models. We find that COVID-19 cases do not have a significant Granger instantaneous causality, at a 95% confidence level, with crimes such as offense ($p=0.42$) and assault ($p=0.97$). On the contrary, a close statistically significant Granger causality was found between COVID-19 cases impacting vehicle theft ($p=0.0009$) and an instantaneous causality was found with thefts ($0.005$). The impulse response of a new case on these crimes and the model predictions are outlined in Figures \@ref(fig:r31) and \@ref(fig:r32).

```{r top 5 crime VAR3, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
# extract top 5 crime
top5crime <- phil %>%
  filter(text_general_code %in% head(phil_summary$text_general_code, 5)) %>%
  group_by(dispatch_date, text_general_code) %>%
  tally() %>%
  spread(text_general_code, n)

# rename columns
colnames(top5crime) <- c('time',
                         "offense",
                         "assault",
                         "vehicle",
                         "thefts",
                         "vandalism")

# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))

# use loaded covid data 
# calculate the difference per day
covid19_PA_diff <- data.frame(diff(covid19_PA$confirmed))
colnames(covid19_PA_diff)[1] = "confirmed"
covid19_PA_diff$date = covid19_PA$date[2:length(covid19_PA$date)]
# extract for tranforming into time series data
ts_PA <- covid19_PA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# plot time series of PA infection
#ts_plot(ts_PA)
# conduct ADF Test
#adf.test(as.ts(ts_PA))
# not stationary!

# try first log difference
ts_diff_PA <- diff(ts_PA)
#ts_plot(ts_diff_PA)
# still clearly not stationary
# need de-trend

# de-trend 
# GAMM model from STA303 A3

# time as integer
covid19_PA_diff$timeInt = as.numeric(covid19_PA_diff$date)
# make a copy to avoid perfect collinearity
covid19_PA_diff$timeIid = covid19_PA_diff$timeInt

# make a copy to avoid perfect collinearity
covid19_PA_diff$timeIid = covid19_PA_diff$timeInt

# GAMM model
# 50 too overfit. 15 looks decent
#changed to 90

# covid19_PA_diff <- covid19_PA_diff %>%
#             filter(confirmed > 0)
covid19_PA_diff$confirmed[covid19_PA_diff$confirmed < 0] <- 0

gamPA <- gamm4::gamm4((confirmed) ~  s(timeInt, k=90), random = ~(1|timeIid), 
	data=covid19_PA_diff, family=poisson(link='log'))

#fitted value
toPredict = data.frame(time = seq(covid19_PA_diff$date[1], 
                                          covid19_PA_diff$date[length(covid19_PA_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# obtain forecast
forecast <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamPA$gam, toPredict, se.fit=TRUE))))

# access residuals
PA_res <- data.frame(covid19_PA_diff$confirmed - forecast$fit)

# transform into time series
PA_res$time = covid19_PA_diff$date
colnames(PA_res)[1] = "residuals"

col_order <- c("time", "residuals")
PA_res <- PA_res[, col_order]

PA_res_ts <- ts_xts(PA_res)

# specify common time range
# start from when covid was a thing
# end with crime since it is manually updated
common_time <- seq.Date(start(PA_res_ts), as.Date("2021-12-31"), by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       PA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
optimal_offense <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_assault <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_vehicle <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_thefts <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

VAR_offense <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]), p=optimal_offense$selection[1])
VAR_assault <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]), p=optimal_assault$selection[1])
VAR_vehicle <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]), p=optimal_vehicle$selection[1])
VAR_thefts <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]), p=optimal_thefts$selection[1])
VAR_vandalism <- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]), p=optimal_vandalism$selection[1])
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
causality(VAR_offense, cause = colnames(combined_diff)[1])
causality(VAR_offense, cause = "residuals")

causality(VAR_assault, cause = colnames(combined_diff)[2])
causality(VAR_assault, cause = "residuals")

causality(VAR_vehicle, cause = colnames(combined_diff)[3])
causality(VAR_vehicle, cause = "residuals")

causality(VAR_thefts, cause = colnames(combined_diff)[4])
causality(VAR_thefts, cause = "residuals")

```

```{r r31, fig.show="hold",  fig.cap = "Impact of a new COVID-19 case on vehicle thefts (left), and VAR model prediction for vehicle thefts (right).", out.width="50%", message = FALSE, echo = FALSE, warning=FALSE}

lags = c(1:16)
par(mfrow = c(1,2))

# general vehicle theft
# irf
irf_vehicle_1 <- irf(VAR_vehicle,
                     impulse = "residuals",
                     response = "vehicle",
                     n.ahead = 15)
# ggplot
irf_vehicle_1_gg <- data.frame(
  irf_vehicle_1$irf$residuals[,1],
  irf_vehicle_1$Lower$residuals[,1],
  irf_vehicle_1$Upper$residuals[,1]
  )

colnames(irf_vehicle_1_gg) <- c("mean", "lower", "upper")

irf_vehicle_1_plot <- ggplot(irf_vehicle_1_gg, aes(x=lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  ggtitle("") +
  xlab("Days after a New COVID-19 Case")+
  ylab("Vehicle Thefts")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

irf_vehicle_1_plot

f_vehicle <- forecast::forecast(VAR_vehicle)

autoplot(f_vehicle$forecast$vehicle) +
  ggtitle("Prediction on how many more vehicle theft cases compared to yesterday") +
  theme_minimal() +
  ylab("Day-to-day change") +
  xlab(paste("Number of days since", common_time[1]))

```

```{r r32, fig.show="hold",  fig.cap = "Impact of a new COVID-19 case on thefts (left), and VAR model prediction for thefts (right).", out.width="50%", message = FALSE, echo = FALSE, warning=FALSE}

# theft
# irf
irf_theft_1 <- irf(VAR_thefts,
                   impulse = "residuals",
                   response = "thefts",
                   n.ahead = 15)

# ggplot
irf_theft_1_gg <- data.frame(
  irf_theft_1$irf$residuals[,1],
  irf_theft_1$Lower$residuals[,1],
  irf_theft_1$Upper$residuals[,1]
)

colnames(irf_theft_1_gg) <- c("mean", "lower", "upper")

irf_theft_1_plot <- ggplot(irf_theft_1_gg, aes(x=lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  ggtitle("") +
  xlab("Days after a new COVID-19 case")+
  ylab("Thefts")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

irf_theft_1_plot

f_theft <- forecast::forecast(VAR_thefts)

autoplot(f_theft$forecast$thefts) +
  ggtitle("Prediction on how many more thefts compared to yesterday") +
  theme_minimal() +
  ylab("Day-to-day change") +
  xlab(paste("Number of days since", common_time[1]))

```

## Seattle

Finally, for the city of Seattle, only burglaries stood out based on the fitted VAR models. We find that COVID-19 cases do not have a significant Granger instantaneous causality, at a 95% confidence level, with crimes such as larceny ($p=0.25$), vandalism ($p=0.79$), vehicle theft ($p=0.31$) and theft from vehicle ($p=0.46$). With a relaxed 90% confidence level, a statistically significant Granger causality was found between COVID-19 cases impacting burglaries ($p=0.073$). The impulse response of a new case on burglaries and the model predictions are outlined in Figure \@ref(fig:r41).

```{r top 5 crime VAR4, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
# extract top 5 crime
top5crime <- seattle %>%
  filter(offense %in% head(seattle_summary$offense, 5)) %>%
  group_by(Date, offense) %>%
  tally() %>%
  spread(offense, n)

# rename columns
colnames(top5crime) <- c('time',
                         "larceny",
                         "burglary",
                         "vandalism",
                         'vehicle_theft',
                         "theft_from_vehicle")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
# plot cumulative cases
# extract for transforming into time series
ts_SEA <- covid19_SEA %>%
  dplyr::select(date, confirmed) %>%
  ts_xts()

# plot daily cases
# first difference
ts_diff_SEA <- na.omit(diff(ts_SEA))

covid19_SEA_diff <- data.frame(diff(covid19_SEA$confirmed) - min(ts_diff_SEA))

colnames(covid19_SEA_diff)[1] = "confirmed"
covid19_SEA_diff$date = covid19_SEA$date[2:length(covid19_SEA$date)]

# time as integer
covid19_SEA_diff$timeInt = as.numeric(covid19_SEA_diff$date)
# RIke a copy to avoid perfect collinearity for mixed effect
covid19_SEA_diff$timeIid = covid19_SEA_diff$timeInt

# GAMM model
gamSEA <- gamm4::gamm4(confirmed ~ s(timeInt, k = 90), 
                      random = ~(1|timeIid),
                      data = covid19_SEA_diff,
                      family = poisson(link = 'log'))

# plot fitted value
toPredict = data.frame(time = seq(covid19_SEA_diff$date[1],
                                  covid19_SEA_diff$date[length(covid19_SEA_diff$date)],
                                  by = '1 day'))

toPredict$timeInt = as.numeric(toPredict$time)
forecast_covid <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamSEA$gam, toPredict, se.fit=TRUE))))

# access residuals
SEA_res <- data.frame(covid19_SEA_diff$confirmed - forecast_covid$fit)

# transform into time series
SEA_res$time = covid19_SEA_diff$date
colnames(SEA_res)[1] = "residuals"

col_order <- c("time", "residuals")
SEA_res <- SEA_res[, col_order]

SEA_res_ts <- ts_xts(SEA_res)

# start from when covid was a thing
# end with 1 day before today's date
common_time <- seq.Date(start(SEA_res_ts), as.Date("2021-12-31") , by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       SEA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
# variable selection based on AIC
optimal_larceny <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_burglary <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_vehicle_theft <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_theft_fromvehicle <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

# use AIC as selection criteria
VAR_larceny <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]),
                   p=optimal_larceny$selection[1])
VAR_burglary <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]),
                     p=optimal_burglary$selection[1])
VAR_vandalism <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                    p=optimal_vandalism$selection[1])
VAR_vehicle_theft <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                         p=optimal_vehicle_theft$selection[1])
VAR_theft_fromvehicle<- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]),
                              p=optimal_theft_fromvehicle$selection[1])
```

```{r, message = FALSE, echo = FALSE, warning=FALSE, include=FALSE}
causality(VAR_larceny, cause = colnames(combined_diff)[1])
causality(VAR_larceny, cause = "residuals")

causality(VAR_burglary, cause = colnames(combined_diff)[2])
causality(VAR_burglary, cause = "residuals")
causality(VAR_vandalism, cause = colnames(combined_diff)[3])
causality(VAR_vandalism, cause = "residuals")
causality(VAR_vehicle_theft, cause = colnames(combined_diff)[4])
causality(VAR_vehicle_theft, cause = "residuals")
causality(VAR_theft_fromvehicle, cause = colnames(combined_diff)[5])
causality(VAR_theft_fromvehicle, cause = "residuals")

```

```{r r41, fig.show="hold",  fig.cap = "Impact of a new COVID-19 case on burglaries (left), and VAR model prediction for burglaries (right). ", out.width="50%", message = FALSE, echo = FALSE, warning=FALSE}
lags = c(1:16)

par(mfrow = c(1,2))
# only covid significant to bulglary
irf_burglary_1 <- irf(VAR_burglary,
                      impulse = "residuals",
                      response = "burglary",
                      n.ahead = 15)
# ggplot
irf_burglary_1_gg <- data.frame(irf_burglary_1$irf$residuals[,1],
                                irf_burglary_1$Lower$residuals[,1],
                                irf_burglary_1$Upper$residuals[,1])

colnames(irf_burglary_1_gg) <- c("mean", "lower", "upper")

irf_burglary_1_plot <- ggplot(irf_burglary_1_gg, aes(x=lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  xlab("Number of days after a confimed COVID-19 case")+
  ylab("Number of bulglary cases")

f_offense <- forecast::forecast(VAR_offense)

irf_burglary_1_plot

f_burglary <- forecast::forecast(VAR_burglary)

autoplot(f_burglary$forecast$burglary) +
  ggtitle("Prediction on how many more burglarys compared to yesterday") +
  theme_minimal() +
  ylab("Day-to-day change") +
  xlab(paste("Number of days since", common_time[1]))

```

Across all of the studied cities, we found that the crimes most significantly associated with COVID-19 cases are vehicle thefts and burglaries/thefts, with vandalism and assault also emerging as significant in some sectors. In the subsequent section we now discuss various implications of these findings and also discuss limitations of the study.

\newpage
# Discussion

Pellentesque id purus eros. Curabitur rutrum ut nisl vel facilisis. Maecenas dapibus vel justo ut aliquet. Suspendisse at auctor sem. Pellentesque sit amet tortor vel arcu sollicitudin lacinia. Pellentesque condimentum, metus eget maximus aliquam, neque nisl.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris ultricies felis in neque hendrerit semper. Sed quis ullamcorper nulla. Donec a mauris ut ipsum interdum tempor a et urna. Ut vehicula ligula sit amet nunc sodales, et facilisis felis vestibulum. Ut ullamcorper erat lorem, id congue ex ultricies ac. Nunc fermentum neque orci. Quisque congue, ligula eget tincidunt condimentum, nulla leo suscipit nibh, eu auctor tellus augue nec turpis. Nulla lorem dui, scelerisque non molestie et, pharetra in enim. Curabitur placerat leo a purus pellentesque, lobortis hendrerit justo posuere. Fusce condimentum porttitor purus, eget ullamcorper elit bibendum et. Cras quam mauris, mollis a cursus non, cursus id neque. Cras eget elit posuere, ornare dui non, posuere neque.

Ut tempor vestibulum augue sit amet hendrerit. Morbi rhoncus dignissim velit. Aliquam erat volutpat. Maecenas semper nulla nec sem tempus, eget elementum odio pulvinar. Donec sit amet nibh in risus dictum venenatis. Mauris non volutpat risus, vitae ornare dui. Praesent blandit leo nec dolor convallis, ut convallis lacus bibendum. Donec nec luctus quam. Nam congue suscipit ex, sed suscipit ligula vestibulum vel. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Duis suscipit sem at enim commodo eleifend. Aenean eu malesuada justo. Sed finibus tincidunt lectus, eget facilisis leo porttitor id.

Morbi pellentesque ligula sed orci lacinia, in vulputate velit vestibulum. Nam commodo lectus odio, nec hendrerit purus consequat et. Sed vel iaculis ex. Maecenas mollis nisi id est pharetra facilisis. Fusce et tempor lorem. Etiam at facilisis nulla. Praesent sed erat odio. Integer sed dictum tellus, in finibus odio. Duis et rhoncus justo.
Pellentesque id purus eros. Curabitur rutrum ut nisl vel facilisis. Maecenas dapibus vel justo ut aliquet. Suspendisse at auctor sem. Pellentesque sit amet tortor vel arcu sollicitudin lacinia. Pellentesque condimentum, metus eget maximus aliquam, neque nisl.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris ultricies felis in neque hendrerit semper. Sed quis ullamcorper nulla. Donec a mauris ut ipsum interdum tempor a et urna. Ut vehicula ligula sit amet nunc sodales, et facilisis felis vestibulum. Ut ullamcorper erat lorem, id congue ex ultricies ac. Nunc fermentum neque orci. Quisque congue, ligula eget tincidunt condimentum, nulla leo suscipit nibh, eu auctor tellus augue nec turpis. Nulla lorem dui, scelerisque non molestie et, pharetra in enim. Curabitur placerat leo a purus pellentesque, lobortis hendrerit justo posuere. Fusce condimentum porttitor purus, eget ullamcorper elit bibendum et. Cras quam mauris, mollis a cursus non, cursus id neque. Cras eget elit posuere, ornare dui non, posuere neque.

Ut tempor vestibulum augue sit amet hendrerit. Morbi rhoncus dignissim velit. Aliquam erat volutpat. Maecenas semper nulla nec sem tempus, eget elementum odio pulvinar. Donec sit amet nibh in risus dictum venenatis. Mauris non volutpat risus, vitae ornare dui. Praesent blandit leo nec dolor convallis, ut convallis lacus bibendum. Donec nec luctus quam. Nam congue suscipit ex, sed suscipit ligula vestibulum vel. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Duis suscipit sem at enim commodo eleifend. Aenean eu malesuada justo. Sed finibus tincidunt lectus, eget facilisis leo porttitor id.

Morbi pellentesque ligula sed orci lacinia, in vulputate velit vestibulum. Nam commodo lectus odio, nec hendrerit purus consequat et. Sed vel iaculis ex. Maecenas mollis nisi id est pharetra facilisis. Fusce et tempor lorem. Etiam at facilisis nulla. Praesent sed erat odio. Integer sed dictum tellus, in finibus odio. Duis et rhoncus justo.
Pellentesque id purus eros. Curabitur rutrum ut nisl vel facilisis. Maecenas dapibus vel justo ut aliquet. Suspendisse at auctor sem. Pellentesque sit amet tortor vel arcu sollicitudin lacinia. Pellentesque condimentum, metus eget maximus aliquam, neque nisl.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris ultricies felis in neque hendrerit semper. Sed quis ullamcorper nulla. Donec a mauris ut ipsum interdum tempor a et urna. Ut vehicula ligula sit amet nunc sodales, et facilisis felis vestibulum. Ut ullamcorper erat lorem, id congue ex ultricies ac. Nunc fermentum neque orci. Quisque congue, ligula eget tincidunt condimentum, nulla leo suscipit nibh, eu auctor tellus augue nec turpis. Nulla lorem dui, scelerisque non molestie et, pharetra in enim. Curabitur placerat leo a purus pellentesque, lobortis hendrerit justo posuere. Fusce condimentum porttitor purus, eget ullamcorper elit bibendum et. Cras quam mauris, mollis a cursus non, cursus id neque. Cras eget elit posuere, ornare dui non, posuere neque.

Ut tempor vestibulum augue sit amet hendrerit. Morbi rhoncus dignissim velit. Aliquam erat volutpat. Maecenas semper nulla nec sem tempus, eget elementum odio pulvinar. Donec sit amet nibh in risus dictum venenatis. Mauris non volutpat risus, vitae ornare dui. Praesent blandit leo nec dolor convallis, ut convallis lacus bibendum. Donec nec luctus quam. Nam congue suscipit ex, sed suscipit ligula vestibulum vel. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Duis suscipit sem at enim commodo eleifend. Aenean eu malesuada justo. Sed finibus tincidunt lectus, eget facilisis leo porttitor id.

Morbi pellentesque ligula sed orci lacinia, in vulputate velit vestibulum. Nam commodo lectus odio, nec hendrerit purus consequat et. Sed vel iaculis ex. Maecenas mollis nisi id est pharetra facilisis. Fusce et tempor lorem. Etiam at facilisis nulla. Praesent sed erat odio. Integer sed dictum tellus, in finibus odio. Duis et rhoncus justo.Pellentesque id purus eros. Curabitur rutrum ut nisl vel facilisis. Maecenas dapibus vel justo ut aliquet. Suspendisse at auctor sem. Pellentesque sit amet tortor vel arcu sollicitudin lacinia. Pellentesque condimentum, metus eget maximus aliquam, neque nisl.

Morbi pellentesque ligula sed orci lacinia, in vulputate velit vestibulum. Nam commodo lectus odio, nec hendrerit purus consequat et. Sed vel iaculis ex. Maecenas mollis nisi id est pharetra facilisis. Fusce et tempor lorem. Etiam at facilisis nulla. Praesent sed erat odio. Integer sed dictum tellus, in finibus odio. Duis et rhoncus justo.Pellentesque id purus eros. Curabitur rutrum ut nisl vel facilisis. Maecenas dapibus vel justo ut aliquet. Suspendisse at auctor sem. Pellentesque sit amet tortor vel arcu sollicitudin lacinia. Pellentesque condimentum, metus eget maximus aliquam, neque nisl.

Morbi pellentesque ligula sed orci lacinia, in vulputate velit vestibulum. Nam commodo lectus odio, nec hendrerit purus consequat et. Sed vel iaculis ex. Maecenas mollis nisi id est pharetra facilisis. Fusce et tempor lorem. Etiam at facilisis nulla. Praesent sed erat odio. Integer sed dictum tellus, in finibus odio. Duis et rhoncus justo.Pellentesque id purus eros. Curabitur rutrum ut nisl vel facilisis. Maecenas dapibus vel justo ut aliquet. Suspendisse at auctor sem. Pellentesque sit amet tortor vel arcu sollicitudin lacinia. Pellentesque condimentum, metus eget maximus aliquam, neque nisl.

\newpage 
<!-- \appendix  -->

<!--  # Appendix {-}  -->
<!--  # Additional details  -->

\newpage

# References